---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---
Reduction Effects upon Dart and Arrow Classification: A New Model Using an
Expanded Dataset
Michael J. Shott, Marcelo Cardillo and Judith Charlin
DOI:https://doi.org/10.5281/zenodo.22968953


Supplementary data

 GENERAL NOTE:
 This R Markdown file contains the complete LDA analysis performed on the
 344-case dataset (LDA_master.txt) to evaluate morphometric classification
 of projectile points as darts or arrows.

 Three Linear Discriminant Analysis (LDA) models are evaluated using different
 combinations of morphometric variables:
   1. Length + Width (LW), following Buchanan et al.
   2. Width + Thickness (WT), following Shott (1997).
   3. Length + Width + Thickness (LWT), to evaluate the contribution of
      Thickness when Length and Width are included.

 The dataset is divided into training and test subsets. Predictor variables
 are standardized using parameters estimated from the training data and the
 same transformation is applied to the test data. Each model is evaluated on
 the independent test set using posterior probabilities and classification
 thresholds of 0.15 and 0.50.

 Model performance is further evaluated using repeated 10-fold cross-validation
 (10 repeats). Accuracy, Kappa, sensitivity, specificity, and balanced accuracy
 are summarized across resampling iterations, including empirical 95% intervals.

 The analysis also examines how classification performance changes across a
 continuous range of posterior-probability thresholds. Additional sections
 identify test specimens whose classification changes across thresholds,
 visualize morphologically ambiguous cases, and explore the relationship
 between misclassification and the Iriarte index (Length / Thickness).

 The script is intended to provide a complete and reproducible record of the
 model fitting, classification, cross-validation, performance evaluation,
 threshold analysis, and examination of misclassified test specimens.




```{r}
#001
# Load required libraries
library(dplyr)
library(caret)
library(ggplot2)
library(MASS)
library(tidyr)
library(purrr)
```


```{r}
#002
# Load the dataset
master_LDA<-read.table("LDA_master.txt", header=T, row.names=1)

```

```{r}
#003
# Inspect the dataset
head(master_LDA)

summary(master_LDA)

table(master_LDA$type, useNA = "ifany")

```


```{r}
#004
# Create training and test sets
set.seed(123)


train_index <- createDataPartition(master_LDA$type,
                                   p = 0.80,
                                   list = FALSE)

train <- master_LDA[train_index, ]
test  <- master_LDA[-train_index, ]

```


```{r}
#005
# Prepare the length and width model


set.seed(2026)

pred_vars <- c("length", "width")

# standardization only with train
pp <- preProcess(train[, pred_vars],
                 method = c("center", "scale"))

train_std <- train
train_std[, pred_vars] <- predict(pp, train[, pred_vars])

# pred to de test set
test_std <- test
test_std[, pred_vars] <- predict(pp, test[, pred_vars])
```



```{r}
#006
# Fit the first LDA model
lda_mod1 <- lda(type ~ length + width,
                data = train_std)

lda_mod1
```


```{r}
#007
# Predict with the first model
pred1 <- predict(lda_mod1, newdata = test_std)

head(pred1$posterior)

```


```{r}
#008
# Extract first-model probabilities
test_std$prob_dart <- pred1$posterior[, "dart"]

```



```{r}
#009
# Apply 0.15 and 0.50 thresholds

test_std$pred015 <- ifelse(test_std$prob_dart >= 0.15,
                           "dart",
                           "arrow")

test_std$pred050 <- ifelse(test_std$prob_dart >= 0.50,
                           "dart",
                           "arrow")
```


```{r}
#010
# Evaluate first-model classifications

cm015 <- confusionMatrix(
  factor(test_std$pred015,
         levels=c("arrow","dart")),
  factor(test_std$type,
         levels=c("arrow","dart")),
  positive="dart"
)

cm050 <- confusionMatrix(
  factor(test_std$pred050,
         levels=c("arrow","dart")),
  factor(test_std$type,
         levels=c("arrow","dart")),
  positive="dart"
)
```



```{r}
#011
# Review first-model results
print(cm015)

print(cm050)
```

```{r}
#012
# Configure repeated cross-validation

set.seed(2026)

ctrl <- trainControl(
  method = "repeatedcv",
  number = 10,          # 10 folds
  repeats = 10,         # repetido 10 veces
  classProbs = TRUE,
  savePredictions = "final"
)
```





```{r}
#013
# Run analysis step
set.seed(2026)

lda_cv1 <- train(
  type ~ length + width,
  data = train,
  method = "lda",
  trControl = ctrl,
  preProcess = c("center", "scale"),
  metric = "Accuracy"
)
```


```{r}
#014
# Review first-model cross-validation
print(lda_cv1)

print(lda_cv1$results)
```
```{r}
#015
# Generate model predictions
#####modelo 2 type ~ width + thickness

set.seed(2026)

# Variables predictoras
pred_vars2 <- c("width", "thickness")

# Ajustar la estandarización SOLO con train
pp2 <- preProcess(train[, pred_vars2],
                 method = c("center", "scale"))

# Aplicarla a train
train_std2 <- train
train_std2[, pred_vars2] <- predict(pp2, train[, pred_vars2])

# Aplicarla a test usando los mismos parámetros
test_std2 <- test
test_std2[, pred_vars2] <- predict(pp2, test[, pred_vars2])

```



```{r}
#016
# Fit the second LDA model
lda_mod2 <- lda(type ~width+thickness,
                data = train_std2)

lda_mod2
```

```{r}
#017
# Generate model predictions
pred2 <- predict(lda_mod2, newdata = test_std2)

head(pred2$posterior)
```


```{r}
#018
# Extract second-model probabilities

test_std2$prob_dart <- pred2$posterior[, "dart"]
```



```{r}
#019
# Apply classification thresholds

test_std2$pred015_2 <- ifelse(test_std2$prob_dart >= 0.15,
                           "dart",
                           "arrow")

test_std2$pred050_2 <- ifelse(test_std2$prob_dart >= 0.50,
                           "dart",
                           "arrow")
```



```{r}
#020
# Evaluate second-model classifications

cm015_2 <- confusionMatrix(
  factor(test_std2$pred015_2,
         levels=c("arrow","dart")),
  factor(test_std2$type,
         levels=c("arrow","dart")),
  positive="dart"
)

cm050_2 <- confusionMatrix(
  factor(test_std2$pred050_2,
         levels=c("arrow","dart")),
  factor(test_std2$type,
         levels=c("arrow","dart")),
  positive="dart"
)
```


```{r}
#021
# Review second-model results

print(cm015_2)

print(cm050_2)

```

```{r}
#022
# 10 fold CV, modelo 2 ##esto es igual

set.seed(2026)

ctrl <- trainControl(
  method = "repeatedcv",
  number = 10,          # 10 folds
  repeats = 10,         # repetido 10 veces
  classProbs = TRUE,
  savePredictions = "final"
)
```




```{r}
#023
# Run analysis step
set.seed(2026)

lda_cv2 <- train(
  type ~ width+thickness,
  data = train,
  method = "lda",
  trControl = ctrl,
  preProcess = c("center", "scale"),
  metric = "Accuracy"
)
```



```{r}
#024
# Review second-model cross-validation
print(lda_cv2)

print(lda_cv2$results)
```


```{r}
#025
# Generate model predictions
#####modelo 3 type ~ length+width + thickness

set.seed(2026)

# Variables predictoras
pred_vars3 <- c( "length", "width", "thickness")

# Ajustar la estandarización SOLO con train
pp3 <- preProcess(train[, pred_vars3],
                 method = c("center", "scale"))

# Aplicarla a train
train_std3 <- train
train_std3[, pred_vars3] <- predict(pp3, train[, pred_vars3])

# Aplicarla a test usando los mismos parámetros
test_std3 <- test
test_std3[, pred_vars3] <- predict(pp3, test[, pred_vars3])


```



```{r}
#026
# Fit the third LDA model

lda_mod3 <- lda(type ~ length + width+thickness,
                data = train_std3)

lda_mod3
```
```{r}
#027
# Generate model predictions
pred3 <- predict(lda_mod3, newdata = test_std3)

head(pred3$posterior)
```


```{r}
#028
# Extract third-model probabilities

test_std3$prob_dart <- pred3$posterior[, "dart"]
```



```{r}
#029
# Apply classification thresholds

test_std3$pred015_3 <- ifelse(test_std3$prob_dart >= 0.15,
                           "dart",
                           "arrow")

test_std3$pred050_3 <- ifelse(test_std3$prob_dart >= 0.50,
                           "dart",
                           "arrow")
```



```{r}
#030
# Evaluate third-model classifications

cm015_3 <- confusionMatrix(
  factor(test_std3$pred015_3,
         levels=c("arrow","dart")),
  factor(test_std3$type,
         levels=c("arrow","dart")),
  positive="dart"
)

cm050_3 <- confusionMatrix(
  factor(test_std3$pred050_3,
         levels=c("arrow","dart")),
  factor(test_std3$type,
         levels=c("arrow","dart")),
  positive="dart"
)
```


```{r}
#031
# Review third-model results

print(cm015_3)

print(cm050_3)

```

```{r}
#032
# 10 fold CV, modelo 3 ##esto es igual

set.seed(2026)

ctrl <- trainControl(
  method = "repeatedcv",
  number = 10,          # 10 folds
  repeats = 10,         # repetido 10 veces
  classProbs = TRUE,
  savePredictions = "final"
)
```




```{r}
#033
# Run analysis step
set.seed(2026)

lda_cv3 <- train(
  type ~ length + width+thickness,
  data = train,
  method = "lda",
  trControl = ctrl,
  preProcess = c("center", "scale"),
  metric = "Accuracy"
)
```



```{r}
#034
# Review third-model cross-validation
print(lda_cv3)

print(lda_cv3$results)
```

```{r}
#035
# Collect the fitted models

modelos <- list(
  m1 = lda_cv1,
  m2 = lda_cv2,
  m3 = lda_cv3
)

```


```{r}
#036
# Calculate cross-validation metrics

extract_metrics <- function(modelo, nombre){

   modelo$pred %>%
    group_by(Resample) %>%
    group_modify(~{

      cm <- confusionMatrix(
        factor(.x$pred,
               levels = c("arrow","dart")),
        factor(.x$obs,
               levels = c("arrow","dart")),
        positive = "dart"
      )

      tibble(
        Accuracy = unname(cm$overall["Accuracy"]),
        Kappa = unname(cm$overall["Kappa"]),
        Sensitivity = unname(cm$byClass["Sensitivity"]),
        Specificity = unname(cm$byClass["Specificity"]),
        BalancedAccuracy = unname(cm$byClass["Balanced Accuracy"])
      )

    }) %>%

    mutate(Model = nombre)
}
```


```{r}
#037
# extraer metricas 3 modelos

m1 <- extract_metrics(lda_cv1, "Length + Width")

m2 <- extract_metrics(lda_cv2, "Width + Thickness")

m3 <- extract_metrics(lda_cv3, "Length + Width + Thickness")
```



```{r}
#038
# Combine model metrics
metricas <- bind_rows(m1, m2, m3)
```


```{r}
#039
# Summarize model performance
#####comparacion modelos intervalos empiricos 95%


##check metricas
metricas %>%
  group_by(Model) %>%
  summarise(
    n = n(),
    Accuracy = mean(Accuracy),
    BA = mean(BalancedAccuracy),
    Sensitivity = mean(Sensitivity),
    Specificity = mean(Specificity),
    Kappa = mean(Kappa)
  )
```

```{r}
#040
# Summarize model performance
metricas_long <- metricas %>%
  pivot_longer(
    cols = c(
      Accuracy,
      BalancedAccuracy,
      Sensitivity,
      Specificity,
      Kappa
    ),
    names_to = "Metric",
    values_to = "Value"
  )
```


```{r}
#041
# Summarize model performance
resumen_metrica <- function(x){

  tibble(
    media = mean(x),
    sd = sd(x),
    mediana = median(x),
    LI95 = quantile(x, 0.025),
    LS95 = quantile(x, 0.975)
  )

}


resumen <- metricas %>%
  pivot_longer(
    cols = c(Accuracy,
             BalancedAccuracy,
             Sensitivity,
             Specificity,
             Kappa),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  group_by(Model, Metric) %>%
  group_modify(~ resumen_metrica(.x$Value))

```



```{r}
#042
# Run analysis step
print(resumen)
```

```{r}
#043
# Plot model performance

ggplot(resumen,
       aes(x = Model,
           y = media)) +

  geom_point(size = 3,
             colour = "black") +

  geom_errorbar(aes(ymin = LI95,
                    ymax = LS95),
                width = .15,
                linewidth = .8) +

  facet_wrap(~Metric,
             ncol = 3,
             scales = "free_y") +

  labs(x = "",
       y = "Mean performance (95% empirical CI)") +

  theme_classic(base_size = 14) +

  theme(
    strip.background = element_rect(fill = "white",
                                    colour = "black"),
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 15,
                               hjust = 1)
  )

```
```{r}
#044
# Summarize model performance
metricas_long$Model <- factor(
  metricas_long$Model,
  levels = c(
    "Length + Width",
    "Width + Thickness",
    "Length + Width + Thickness"
  ),
  labels = c("LW", "WT", "LWT")
)
```


```{r}
#045
# Visualize analysis results
ggplot(metricas_long,
       aes(x = Model,
           y = Value)) +

  geom_jitter(width = 0.08,
              alpha = 0.35,
              size = 1.8,
              colour = "grey45") +

  stat_summary(fun = median,
               geom = "point",
               shape = 15,
               size = 3.5,
               colour = "black") +

  facet_wrap(~Metric,
             ncol = 3,
             scales = "free_y") +

  labs(x = "",
       y = "Performance metric") +

  theme_classic(base_size = 14) +

  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white",
                                    colour = "black"),
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 0,
                               hjust = 1)
  )
```
```{r}
#046
# Evaluate classification performance


threshold_curve <- function(modelo,
                            nombre,
                            thresholds = seq(0.05,0.95,0.05)){

  resultados <- list()

  for(th in thresholds){

    tmp <- modelo$pred %>%

      group_by(Resample) %>%

      group_modify(~{

        pred <- factor(
          ifelse(.x$dart >= th,
                 "dart",
                 "arrow"),
          levels=c("arrow","dart")
        )

        cm <- confusionMatrix(
          pred,
          factor(.x$obs,
                 levels=c("arrow","dart")),
          positive="dart"
        )

        tibble(

          Sensitivity = unname(cm$byClass["Sensitivity"]),

          Specificity = unname(cm$byClass["Specificity"]),

          BalancedAccuracy =
            unname(cm$byClass["Balanced Accuracy"])

        )

      }) %>%

      mutate(

        Threshold = th,

        Model = nombre

      )

    resultados[[length(resultados)+1]] <- tmp

  }

  bind_rows(resultados)

}
```


```{r}
#047
# Analyze classification thresholds
thr1 <- threshold_curve(lda_cv1,"LW")

thr2 <- threshold_curve(lda_cv2,"WT")

thr3 <- threshold_curve(lda_cv3,"LWT")
```


```{r}
#048
# Run analysis step
thresholds <- bind_rows(
  thr1,
  thr2,
  thr3
)
```


```{r}
#049
# 95%
thr_summary <-

thresholds %>%

pivot_longer(

cols=c(
Sensitivity,
Specificity,
BalancedAccuracy),

names_to="Metric",

values_to="Value"

) %>%

group_by(

Model,

Metric,

Threshold

) %>%

summarise(

Mean=mean(Value),

LI95=quantile(Value,.025),

LS95=quantile(Value,.975),

.groups="drop"

)
```




```{r}
#050
# Visualize analysis results

ggplot(thr_summary,
       aes(x = Threshold,
           y = Mean,
           colour = Model)) +

  geom_line(linewidth = 1.1) +

  ## Thresholds discutidos por Buchanan
  geom_vline(xintercept = c(0.15, 0.50),
             linetype = "dashed",
             colour = "grey40",
             linewidth = 0.6) +

  facet_wrap(~Metric,
             ncol = 1) +

  coord_cartesian(ylim = c(0.65, 1.00)) +

  scale_x_continuous(
    breaks = seq(0.1, 0.9, 0.1)
  ) +

  labs(
    x = "Classification threshold",
    y = "Mean performance",
    colour = "Model"
  ) +

  theme_classic(base_size = 14) +

  theme(
    legend.position = "top",
    legend.title = element_text(face = "bold"),
    strip.background = element_rect(fill = "white",
                                    colour = "black"),
    strip.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold")
  )
```


```{r}
#051
# Build the classification table
tabla <- data.frame(
  ID   = rownames(test),
  Real = test$type,
  stringsAsFactors = FALSE
)

head(tabla)
```


```{r}
#052
# Classification threshold = 0.15
tabla$LW015 <- ifelse(
  pred1$posterior[, "dart"] >= 0.15,
  "dart",
  "arrow"
)

## Classification threshold = 0.50
tabla$LW050 <- ifelse(
  pred1$posterior[, "dart"] >= 0.50,
  "dart",
  "arrow"
)

## P (Dart)
tabla$Prob_LW <- tabla$Prob_LW <- round(pred1$posterior[, "dart"], 4)

## Missclasification rate
tabla$LW015_OK <- ifelse(
  tabla$LW015 == tabla$Real,
  "✓",
  "✗"
)

tabla$LW050_OK <- ifelse(
  tabla$LW050 == tabla$Real,
  "✓",
  "✗"
)
```

```{r}
#053
# Run analysis step
tabla$LW_Cambio <- ifelse(tabla$LW015 != tabla$LW050,
                          "Yes",
                          "No")

```



```{r}
#054
# Run analysis step
tabla$N_aciertos <-
  rowSums(tabla[, c("LW015_OK",
                    "LW050_OK")] == "✓")
```





```{r}
#055
# Run analysis step
## MODEL 2 (Width + Thickness)
##==========================

## Threshold = 0.15
tabla$WT015 <- ifelse(
  pred2$posterior[, "dart"] >= 0.15,
  "dart",
  "arrow"
)

## Threshold = 0.50
tabla$WT050 <- ifelse(
  pred2$posterior[, "dart"] >= 0.50,
  "dart",
  "arrow"
)

## Posterior prob
tabla$Prob_WT <- signif(
  pred2$posterior[, "dart"],
  3
)

## Correct / Incorrec
tabla$WT015_OK <- ifelse(
  tabla$WT015 == tabla$Real,
  "✓",
  "✗"
)

tabla$WT050_OK <- ifelse(
  tabla$WT050 == tabla$Real,
  "✓",
  "✗"
)

## table
tabla$WT_Cambio <- ifelse(
  tabla$WT015 != tabla$WT050,
  "Sí",
  "No"
)
```


```{r}
#056
# Run analysis step
## MODEL 3 (Length + Width + Thickness)
##==========================

tabla$LWT015 <- ifelse(
  pred3$posterior[, "dart"] >= 0.15,
  "dart",
  "arrow"
)

tabla$LWT050 <- ifelse(
  pred3$posterior[, "dart"] >= 0.50,
  "dart",
  "arrow"
)

tabla$Prob_LWT <- signif(
  pred3$posterior[, "dart"],
  3
)

tabla$LWT015_OK <- ifelse(
  tabla$LWT015 == tabla$Real,
  "✓",
  "✗"
)

tabla$LWT050_OK <- ifelse(
  tabla$LWT050 == tabla$Real,
  "✓",
  "✗"
)

tabla$LWT_Cambio <- ifelse(
  tabla$LWT015 != tabla$LWT050,
  "Sí",
  "No"
)
```



```{r}
#057
#   summary

tabla$N_aciertos <- rowSums(
  tabla[, c(
    "LW015_OK","LW050_OK",
    "WT015_OK","WT050_OK",
    "LWT015_OK","LWT050_OK"
  )] == "✓"
)

tabla$N_errores <- 6 - tabla$N_aciertos
```


```{r}
#058
# Run analysis step
tabla$Categoria <- cut(
  tabla$N_errores,
  breaks = c(-1, 0, 5, 6),
  labels = c("Always correct",
             "Parcial errors",
             "Always incorrect")
)
```


```{r}
#059
# Run analysis step
medidas <- data.frame(
  ID = rownames(master_LDA),
  length = master_LDA$length,
  width = master_LDA$width,
  thickness = master_LDA$thickness
)

tabla_plot <- merge(tabla, medidas, by = "ID")
```




```{r}
#060
# Visualize analysis results


ggplot(tabla_plot,
       aes(x = length,
           y = width,
           color = Categoria)) +

  geom_point(size = 3, alpha = 0.8) +

  theme_classic(base_size = 14) +

  labs(
    x = "Length (mm)",
    y = "Width (mm)",
    color = ""
  )
```


```{r}
#061
# Run analysis step
tabla$length    <- test$length
tabla$width     <- test$width
tabla$thickness <- test$thickness
```


```{r}
#062
# Run analysis step
tabla$Grupo <- cut(
  tabla$N_errores,
  breaks = c(-1, 0, 2, 6),
  labels = c("0", "1–2", "3–6")
)
```


```{r}
#063
# Visualize analysis results
ggplot(tabla,
       aes(x = length,
           y = width,
           colour = Categoria,
           shape = Real)) +

  geom_point(size = 3) +

  geom_text(
    data = subset(tabla, N_errores > 0),
    aes(label = ID),
    nudge_y = 0.4,
    size = 3.5,
    show.legend = FALSE
  ) +

  scale_shape_manual(values = c(16,17)) +

  scale_colour_manual(
    values = c(
      "Always correct" = "grey60",
      "Parcial errors" = "green",
      "Always incorrect" = "red"
    )
  ) +

  theme_classic(base_size = 14) +

  labs(
    x = "Length (mm)",
    y = "Width (mm)",
    colour = "Classification",
    shape = "Tipo"
  )
```

```{r}
#064
# List misclassified specimens

errores <- subset(tabla, N_errores > 0)

errores
```




```{r}
#065
# Run analysis step
errores <- errores[
  order(-errores$N_errores),
  c("ID",
    "Real",
    "length",
    "width",
    "thickness",
    "N_errores",
    "Prob_LW",
    "Prob_WT",
    "Prob_LWT",
    "LW015",
    "WT015",
    "LWT015")
]

errores
```


```{r}
#066
# Calculate the Iriarte index

tabla$Iriarte <- tabla$length / tabla$thickness
```



```{r}
#067
# Run analysis step
dart <- subset(tabla, Real == "dart")

dart$Misclassified <- ifelse(
  dart$LW015 == "arrow",
  1,0)

table(dart$Misclassified)
```


```{r}
#068
# Fit logistic regression
mod_log <- glm(
  Misclassified ~ Iriarte,
  family = binomial,
  data = dart
)

summary(mod_log)
```
>mayor indice de iriarte disminuye el odds de missclass 


```{r}
#069
# Run analysis step
exp(coef(mod_log))
```

```{r}
#070
# Run analysis step
exp(confint(mod_log))
```

```{r}
#071
# Run analysis step
anova(mod_log,
      test="Chisq")
```



```{r}
#072
# Run analysis step
reduccion_error=(0.4149904-1)

reduccion_error

```



```{r}
#073
# Visualize analysis results

ggplot(dart,
       aes(Iriarte,
           Misclassified)) +

  geom_jitter(
    height = 0.05,
    width = 0
  ) +

  stat_smooth(
    method = "glm",
    method.args = list(family = "binomial"),
    se = TRUE
  ) +

  theme_classic(base_size = 14) +

  labs(
    x = "Iriarte index",
    y = "Probability of misclassification"
  )
```

```{r}
#074
# Load required packages
library(pscl)

pR2(mod_log)
```



```{r}
#075
# Load required packages
library(pROC)

roc_obj <- roc(
  dart$Misclassified,
  fitted(mod_log)
)

auc(roc_obj)
```




```{r}
#076
# Generate model predictions
predict(
  mod_log,
  newdata=data.frame(Iriarte=c(6,12)),
  type="response"
)
```

```{r}
#077
# Run analysis step
range(dart$Iriarte)
```

```{r}
#078
# Visualize analysis results
library(ggplot2)
library(patchwork)

## Predicciones
newdat <- data.frame(
  Iriarte = seq(min(dart$Iriarte),
                max(dart$Iriarte),
                length.out = 300)
)

pred <- predict(mod_log,
                newdata = newdat,
                type = "link",
                se.fit = TRUE)

newdat$fit <- plogis(pred$fit)
newdat$lwr <- plogis(pred$fit - 1.96*pred$se.fit)
newdat$upr <- plogis(pred$fit + 1.96*pred$se.fit)

##==========================
## Panel A
##==========================

p1 <-

ggplot(newdat,
       aes(Iriarte, fit))+

geom_ribbon(aes(ymin=lwr,
                ymax=upr),
            fill="grey80",
            alpha=.6)+

geom_line(linewidth=1.3,
          colour="#2166AC")+

theme_classic(base_size=15)+

coord_cartesian(ylim=c(0,1))+

labs(x="",
     y="Predicted probability\nof misclassification")

##==========================
## Panel B
##==========================

p2 <-

ggplot(dart,
       aes(Iriarte,
           factor(Misclassified)))+

geom_rug(aes(colour=factor(Misclassified)),
         sides="b",
         linewidth=2)+

geom_point(aes(colour=factor(Misclassified)),
           size=3,
           alpha=.8,
           position=position_jitter(height=.05))+

geom_text(
  data=subset(dart,Misclassified==1),
  aes(label=ID),
  nudge_y=.15,
  show.legend=FALSE,
  colour="red",
  size=4
)+

scale_y_discrete(
  labels=c("Correct","Misclassified")
)+

scale_colour_manual(
  values=c("black","red"),
  guide="none"
)+

theme_classic(base_size=15)+

labs(x="Iriarte index (Length / Thickness)",
     y="Observed")

##==========================
## Final plot
##==========================

p1/p2 +
plot_annotation(tag_levels="A")
```
```{r}
#079
# Run analysis step
## WT
##=====================================

dart <- subset(tabla, Real == "dart")

dart$Misclassified <- ifelse(
  dart$WT015 == "arrow",
  1,0
)

table(dart$Misclassified)
```





```{r}
#080
# Fit logistic regression
mod_log_WT <- glm(
  Misclassified ~ Iriarte,
  family = binomial,
  data = dart
)

summary(mod_log_WT)
```


```{r}
#081
# Run analysis step
exp(coef(mod_log_WT))

exp(confint(mod_log_WT))
```


```{r}
#082
# Run analysis step
anova(
  mod_log_WT,
  test="Chisq"
)
```





```{r}
#083
# Run analysis step
roc_obj_WT <- roc(
  dart$Misclassified,
  fitted(mod_log_WT)
)

auc(roc_obj_WT)
```

```{r}
#084
# Generate model predictions
predict(
  mod_log_WT,
  newdata=data.frame(
    Iriarte=c(6,12)
  ),
  type="response"
)
```


```{r}
#085
# Generate model predictions
newdat <- data.frame(
  Iriarte=seq(
    min(dart$Iriarte),
    max(dart$Iriarte),
    length.out=300
  )
)

pred <- predict(
  mod_log_WT,
  newdata=newdat,
  type="link",
  se.fit=TRUE
)

newdat$fit <- plogis(pred$fit)
newdat$lwr <- plogis(pred$fit-1.96*pred$se.fit)
newdat$upr <- plogis(pred$fit+1.96*pred$se.fit)
```



```{r}
#086
# Visualize analysis results

## Panel A
##==========================

p1 <-

ggplot(newdat,
       aes(Iriarte, fit))+

geom_ribbon(aes(ymin=lwr,
                ymax=upr),
            fill="grey80",
            alpha=.6)+

geom_line(linewidth=1.3,
          colour="#2166AC")+

theme_classic(base_size=15)+

coord_cartesian(ylim=c(0,1))+

labs(x="",
     y="Predicted probability\nof misclassification")

##==========================
## Panel B
##==========================

p2 <-

ggplot(dart,
       aes(Iriarte,
           factor(Misclassified)))+

geom_rug(aes(colour=factor(Misclassified)),
         sides="b",
         linewidth=2)+

geom_point(aes(colour=factor(Misclassified)),
           size=3,
           alpha=.8,
           position=position_jitter(height=.05))+

geom_text(
  data=subset(dart,Misclassified==1),
  aes(label=ID),
  nudge_y=.15,
  show.legend=FALSE,
  colour="red",
  size=4
)+

scale_y_discrete(
  labels=c("Correct","Misclassified")
)+

scale_colour_manual(
  values=c("black","red"),
  guide="none"
)+

theme_classic(base_size=15)+

labs(x="Iriarte index (Length / Thickness)",
     y="Observed")

##==========================
## Final plot
##==========================

p1/p2 +
plot_annotation(tag_levels="A")
```



```{r}
#088
# Run analysis step
## LWT
##=====================================

dart <- subset(tabla, Real == "dart")

dart$Misclassified <- ifelse(
  dart$LWT015 == "arrow",
  1,0
)

table(dart$Misclassified)
```

```{r}
#089
# Logit model LWT
mod_log_LWT <- glm(
  Misclassified ~ Iriarte,
  family = binomial,
  data = dart
)

summary(mod_log_LWT)
```


```{r}
#090
# ODD ratios
exp(coef(mod_log_LWT))

exp(confint(mod_log_LWT))

```


```{r}
#091
# Run the likelihood-ratio test
anova(
  mod_log_LWT,
  test = "Chisq"
)
```




```{r}
#092
# Calculate pseudo-R-squared
pR2(mod_log_LWT)
```

```{r}
#093
# Evaluate the ROC curve
roc_obj_LWT <- roc(
  dart$Misclassified,
  fitted(mod_log_LWT)
)

auc(roc_obj_LWT)
```



```{r}
#094
# Generate predicted probabilities
predict(
  mod_log_LWT,
  newdata = data.frame(Iriarte = c(6,12)),
  type = "response"
)
```


```{r}
#095
# Prepare regression plot data


newdat <- data.frame(
  Iriarte = seq(min(dart$Iriarte),
                max(dart$Iriarte),
                length.out = 300)
)

pred <- predict(
  mod_log_LWT,
  newdata = newdat,
  type = "link",
  se.fit = TRUE
)

newdat$fit <- plogis(pred$fit)
newdat$lwr <- plogis(pred$fit - 1.96 * pred$se.fit)
newdat$upr <- plogis(pred$fit + 1.96 * pred$se.fit)
```



```{r}
#096
# Visualize analysis results

## Panel A
##==========================

p1 <-

ggplot(newdat,
       aes(Iriarte, fit))+

geom_ribbon(aes(ymin=lwr,
                ymax=upr),
            fill="grey80",
            alpha=.6)+

geom_line(linewidth=1.3,
          colour="#2166AC")+

theme_classic(base_size=15)+

coord_cartesian(ylim=c(0,1))+

labs(x="",
     y="Predicted probability\nof misclassification")

##==========================
## Panel B
##==========================

p2 <-

ggplot(dart,
       aes(Iriarte,
           factor(Misclassified)))+

geom_rug(aes(colour=factor(Misclassified)),
         sides="b",
         linewidth=2)+

geom_point(aes(colour=factor(Misclassified)),
           size=3,
           alpha=.8,
           position=position_jitter(height=.05))+

geom_text(
  data = subset(dart, Misclassified == "Misclassified"),
  aes(label=ID),
  nudge_y=.15,
  show.legend=FALSE,
  colour="red",
  size=4
)+

scale_y_discrete(
  labels=c("Correct","Misclassified")
)+

scale_colour_manual(
  values=c("black","red"),
  guide="none"
)+

theme_classic(base_size=15)+

labs(x="Iriarte index (Length / Thickness)",
     y="Observed")

##==========================
## Final plot
##==========================

p1/p2 +
plot_annotation(tag_levels="A")
```


```{r}
#097
# Record R session information
sessionInfo()
```


















